$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-04-30 paper/paper.tex
Provide example derivation

@@

Shared, gradual and linear multiplicity types can all be defined as an instance of the properties in Figure \ref{fig:multiplicities}.
For examples of how these multiplicity types are used in typing derivations, refer to \S \ref{example-derivations}.

@@

  The base case $\constr{zero}$ asserts that $t$ is the most immediate type in typing context, and splits the most immediate usage annotation $x$ in the input context into the usage annotation $y$ the usage annotation $z$ in the output context.
  This is the only place in the typing system where the algebra on multiplicities is used.

@@

  The inductive case $\constr{suc}$ appends an extra type to the typing context, and an extra usage annotation to both the input and output usage context.

@@

    \inferrule
    {\opsquared{x}{y}{z}}
    {\constr{zero} : \contains{\gamma \comma t}{\Gamma \comma x}{\constr{zero}}{t}{y}{\Gamma \comma z}}

    \inferrule
    {loc_i : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
    {\constr{suc} \; loc_i : \contains{\gamma \comma t'}{\Gamma \comma x'}{\suc i}{t}{x}{\Delta \comma x'}}
  \end{mathpar}
  \caption{Variable reference judgments.}
  \label{fig:variable-references}
  \end{figure}
\end{nidefinition}

\begin{example}
\todo{example}
\end{example}

@@

  Parallel composition is only concerned with usage contexts.
  It uses the leftovers of the left-hand process to type the right-hand process.

@@

    \inferrule
    {l \hspace{0.3em} : \types{\gamma}{\Gamma \hspace{0.2em}}{P}{\Delta} \\\\
     r : \types{\gamma}{\Delta}{Q}{\Xi}}
    {\constr{comp} \; l \; r : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
  \end{mathpar}
  \caption{Leftover typing for a resource-aware typing system.}
  \label{fig:types}
  \end{figure}
\end{nidefinition}

\subsubsection{Example derivations}
\label{example-derivations}

\todo{OD: in addition to the above, could we have examples of typing a process in each of the above type systems? It can be the same process everywhere and we can show how the typing changes...}

\begin{nilemma}
  \label{lm:types-unused}
  If $P$ is well-typed under input context $\Gamma$ and output context $\Xi$, and variable $i$ is unused within $P$, then $\Gamma_i \equiv \Xi_i$.
\end{nilemma}
\begin{proof}
  By induction on \textsc{Process} and \textsc{Var}.
\end{proof}

\begin{nilemma}
  \label{lm:types-op}
  For any well typed-process $P$ with input context $\Gamma$ and output context $\Xi$, there exists some $\Delta$ such that $\opctx{\Gamma}{\Delta}{\Xi}$.
\end{nilemma}
\begin{proof}
  By induction on \textsc{Process} and \textsc{Var}.
\end{proof}

>>>>

% Example code
\usepackage{listings}

@@

Shared, gradual and linear multiplicity types can all be defined as an instance of the properties in Figure \ref{fig:multiplicities}.
For examples of how these multiplicity types are used in typing derivations, refer to Example \ref{example-derivations}.

@@

  The base case $\constr{here}$ asserts that $t$ is the most immediate type in typing context, and splits the most immediate usage annotation $x$ in the input context into the usage annotation $y$ the usage annotation $z$ in the output context.
  This is the only place in the typing system where the algebra on multiplicities is used.

@@

  The inductive case $\constr{there}$ appends an extra type to the typing context, and an extra usage annotation to both the input and output usage context.

@@

    \inferrule
    {\opsquared{x}{y}{z}}
    {\constr{here} : \contains{\gamma \comma t}{\Gamma \comma x}{\constr{zero}}{t}{y}{\Gamma \comma z}}

    \inferrule
    {loc_i : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
    {\constr{there} \; loc_i : \contains{\gamma \comma t'}{\Gamma \comma x'}{\suc i}{t}{x}{\Delta \comma x'}}
  \end{mathpar}
  \caption{Variable reference judgments.}
  \label{fig:variable-references}
  \end{figure}
\end{nidefinition}

\begin{example}
  The following is an example that asserts that the input usage context is capable of $\li$ on a $\channel{\unit}{\li}$ type at index $\constr{suc} \; \constr{zero}$, and that consuming such $\li$ results in the stated leftovers.
  \begin{lstlisting}[mathescape,caption=Here \_ introduces an anonymous declaration that must be immediately followed by its definition.]
  _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\constr{suc} \; \constr{zero}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
  _ = $\constr{there} \; \constr{here}$
  \end{lstlisting}
\end{example}

@@

@@

    \inferrule
    {l \hspace{0.3em} : \types{\gamma}{\Gamma \hspace{0.2em}}{P}{\Delta} \\\\
     r : \types{\gamma}{\Delta}{Q}{\Xi}}
    {\constr{comp} \; l \; r : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
  \end{mathpar}
  \caption{Leftover typing for a resource-aware typing system.}
  \label{fig:types}
  \end{figure}

  Parallel composition is only concerned with usage contexts.
  It uses the leftovers of the left-hand process to type the right-hand process.
\end{nidefinition}

\begin{example}
  \label{example-derivations}
  The following example uses both linear and shared types to type the process $p$. 
  We use $\omega$ as a synonym for a $(\one \comma \one)$ multiplicity in an algebra for shared types.
  The usage context $\epsilon$ signifies the empty context where all multiplicities are $(\zero \comma \zero)$.
  The typing derivation creates a new channel over which $\li$ multiplicities of a type $\channel{\unit}{\omega}$ are sent.
  This channel is initialized with a linear multiplicity of $(\one \comma \one)$, and therefore must be used exactly once for sending and once for receiving.
  The right-hand parallel subprocess then creates a new channel as per the specification $\channel{\unit}{\omega}$.

  \begin{lstlisting}[mathescape]
  $p$ : $\Process_1$
  $p$ = $\new{} (\comp{\recv{\constr{zero}} (\recv{\constr{zero}} \PO)}
                    {\new{} (\send{\constr{suc} \; \constr{zero}} {\constr{zero}} (\send{\constr{zero}}{\constr{suc} \; \constr{suc} \; \constr{zero}}) \; \PO)})$

  _ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \omega}{p}{\epsilon}$
  _ = $\constr{chan} \; \channel{\unit}{\omega} \; \li \; \one$
      $(\constr{comp} \;
       (\constr{recv} \; \constr{here} \; (\constr{recv} \; \constr{here} \; \constr{end}))$
           $(\constr{chan} \; \unit \; \omega \; \one \; (\constr{send} \; (\constr{there} \; \constr{here}) \; \constr{here} \; (\constr{send} \; \constr{here} \; (\constr{there} \; \constr{here} \; \constr{there}) \; \constr{end})))
       )$
  \end{lstlisting}
\end{example}

\begin{nilemma}
  \label{lm:types-unused}
  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
\end{nilemma}
\begin{proof}
  By induction on \textsc{Process} and \textsc{Var}.
\end{proof}

\begin{nilemma}
  \label{lm:types-op}
  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, there exists some $\Delta$ such that $\opctx{\Gamma}{\Delta}{\Xi}$.
\end{nilemma}
\begin{proof}
  By induction on \textsc{Process} and \textsc{Var}.
\end{proof}
